---
title: "Homework 3"
author: "Lowell Bander"
date: "November 11, 2015"
output: pdf_document
---

# Problem 2, PCA

```{r}
run_PCA <- function() {
  xtrain <- read.csv("~/Desktop/ADJCLOSE.csv",sep = ",")
  xtrain <- xtrain[complete.cases(xtrain),]
  myPCA = prcomp(xtrain, center = TRUE, scale. = TRUE, retx=TRUE);
  print(summary(myPCA))
}

run_PCA()

```

We can therefore see that the top 5 components capture 0.5924 + 0.2092 + 0.1179 + 0.03165 + 0.02779 = 0.97894 = 97.894% of the variance in the data.

```{r}
xtrain <- read.csv("~/Desktop/ADJCLOSE.csv",sep = ",")
xtrain <- xtrain[complete.cases(xtrain),]
myPCA = prcomp(xtrain, center = TRUE, scale. = TRUE, retx=TRUE);
loadings <- myPCA$x
top5 <- subset(loadings, select=-c(PC6,PC7,PC8,PC9,PC10))
```

Didn't finish this last part yet.

# Problem 3, Random Matrix Theory

We can compute and plot a histogram of the eigenvalues of the empirical covariance matrix like so:

```{r}
load("~/Downloads/CorrMtx_2012_2015_SP500.data")

drawHist <- function(data) {
  empiricalCovarianceMatrix <- cor(data)
  eigenValues <- eigen(empiricalCovarianceMatrix)$values
  hist(eigenValues, breaks=50)  
}

drawHist(RETS)
```

If we now plot the same spectrum, but leave out the largest eigen values, we obtain the following histogram:

```{r}
drawHistLeaveOutLargestEigenValues <- function(data) {
  empiricalCovarianceMatrix <- cor(data)
  eigenValues <- eigen(empiricalCovarianceMatrix)$values
  cleanedEigenValues <- eigenValues[4:length(eigenValues)]
  hist(cleanedEigenValues, breaks=50)  
}

drawHistLeaveOutLargestEigenValues(RETS)
```

From comparing these two histograms, we observe that the largest eigen values are much larger than the remaining eigen values to the point that they dwarf the remaining eigen values. They are so much larger that the standard resolution for the histogram essentially hides the smaller eigen values from view.

```{r}
eigenValuesMatrix = matrix(nrow=472, ncol=50)

for (i in 1:50) {
  RETSRAND = matrix( sample( c(RETS) ) , nrow = dim(RETS)[1], ncol = dim(RETS)[2])
  empiricalCovarianceMatrix <- cor(RETSRAND)
  eigenValues <- eigen(empiricalCovarianceMatrix)$values
  eigenValuesMatrix[,i] <- eigenValues
}

averagedEigens <- rowMeans(eigenValuesMatrix)
hist(averagedEigens, breaks=50)  
```



