---
title: "Midterm"
author: "Lowell Bander"
date: "November 14, 2015"
output: pdf_document
---

# Problem 3

```{r}
library("ISLR", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
plot(Auto)

myAuto <- Auto
myAuto$name <- NULL
cor(myAuto)

model <- lm(mpg~., data = myAuto)
summary(model)
```

We can see from the high R-squared value of 0.8215 that there is a relationship between the predictors and the response. However, _weight_, _year_, _origin_ have the most statistically significant relationship to the response, and displacement somewhat less so.

The coefficient for the _year_ variable suggests that for every year that passes, MPG increases by 0.750773 miles per gallon.

```{r}
plot(model)

m<-lm(mpg ~ cylinders + displacement + horsepower + weight * year + origin, 
   data = myAuto)
summary(m)
```

The interaction between weight and year appears to be statistically significant.

# Problem 3

```{r}
load(file = '~/Downloads/midtermDistanceMatrix.data')

# B = -(1/2)HDH

n <- nrow(DIST)
I <- diag(n)
e <- diag(I)
e_t <- t(e)
H = I - (1/n) * e %*% e_t
B <- -(1/2) * H %*% DIST %*% H

barplot(head(eigen(B)$values, n = 10))
```

We observe that successive eigen values exponentially decay in their representation of the variation in the data.

```{r}
svd <- svd(B)
```
